---
title: "assigmnet3"
author: "QUANG HUY NGUYEN - 45768161"
date: "29/10/2021"
output: pdf_document
editor_options: 
  markdown: 
    wrap: 72
---

# Question 1

a)  

```{r}
library(readxl)
magnes <- read_excel("magnes.xls", col_types = c("numeric", "numeric", "numeric", "skip", "numeric", "numeric", "numeric"))
magnes <- data.frame(Magnesium = c(magnes[[1]],magnes[[4]]),Time = c(magnes[[2]],magnes[[5]]),Treatment = c(magnes[[3]],magnes[[6]]))
magnes$Treatment = as.factor(magnes$Treatment)
magnes$Time = as.numeric(magnes$Time)
head(magnes)
```

```{r}
fit = lm(Magnesium ~ Time + I(Time^2) + Treatment, data = magnes)
summary(fit)
```

-   The model's R-squared is 0.794, which means 79% observation of
    Magnesium can be explain by 3 above variables.
-   Time\^2 is not significant.

```{r, fig.width = 10, fig.height = 10}
par(mfrow = c(2, 2))
plot(fit)
```

-   Linear assumption is not satisfied.

-   Evident of Heteroscedasticity.

-   Q-Q plot show normality.

-   There is 1 extreme points (55) that is exceed 4 standard deviations.

b)  When combine two given models, we have: $$
    E(y_{(56 \times 1)})=X_{(56 \times 3)}\beta_{(1 \times 3)}
    $$ With: $$ 
    X= \begin{bmatrix} 
    1&0&x_1^1&0&x^3_1&0 \\ 
    \vdots&\vdots&\vdots&\vdots&\vdots&\vdots \\
    1&0&x_{56}^1&0&x^3_{56}&0 \\
    0&1&0&x_1^2&0&x^4_1\\ 
    \vdots&\vdots&\vdots&\vdots&\vdots&\vdots \\
    0&1&0&x_{56}^2&0&x^4_{56} \\
    \end{bmatrix}
    $$ And $$ 
    \beta= \begin{bmatrix} \beta_0 \\ \gamma_0 \\ \beta_1 \\ \gamma_1 \\ \beta_2 \\ \gamma_2 \end{bmatrix}
    $$ Assume that: $$
    y_{n×1} \sim N_{56}(\mu,\sigma^2I_n)
    $$ With $$
    \mu=X\beta.
    $$

$$
H_0 : C\beta_{(1 \times 6)} = 0
$$ With $$
C = \begin{bmatrix} 0&1&0 \\ 0&0&1 \end{bmatrix} 
$$ c) T2 test:

```{r}
fit1 = lm(Magnesium ~ Time + I(Time^2), data = magnes)
b <- fit1$coef
X <- model.matrix(fit1)
C <- rbind(c(0,1,0), c(0,0,1))
sig2 <- sum(fit1$res^2)/(dim(magnes)[1]-3)
V <- sig2*C%*%solve(t(X)%*%X)%*%t(C)
d<-t(C%*%b)%*%solve(V)%*%C%*%b
d
cv <- 2*qf(.95, 2, dim(magnes)[1]-3)
cv
```

-   The T value is greater than critical value. Hence, reject the
    hypothesis.

d)  

```{r}
rssf <-sum(fit1$res^2)

fit2 <- lm(Magnesium ~ Time + I(Time^2), data = magnes[28:56,])
rssr <- sum(fit2$residuals^2)
# test statistic:
f = ((rssr - rssf)/(3-1)) / (rssf/(dim(magnes)[1]-3))
f
cv <- 2*qf(.95, 2, dim(magnes)[1]-3)
cv
```

# Question 2

a)  

```{r}
air <- read_excel("air.xls", col_types = c("numeric", 
    "numeric", "numeric", "numeric", "numeric", 
    "numeric", "numeric"))
R <- cor(air)
R
```

```{r}
er <- eigen(R)
v <- er$values
id <- er$vectors
v
cumsum(v)/sum(v)
```

-   The first values will explain \~80% of the total variability. Hence
    select 4 common factors.

c)  

```{r}
S <- var(air)
#assume m = 2
#v[c(1,2)]/sum(v) #variance explained by two factors
Lam = id[, c(1,2)]%*%diag(sqrt(v[c(1,2)])) #factor loading matrix lambda 
Psi = diag(S-Lam%*%t(Lam)) #specific variances
resi = S-(Lam%*%t(Lam)+diag(Psi)) #residual matrix
resi
Psi
```

d)  

```{r}
air.mle <- factanal(air, factors=2) #mle factor analysis
lam.mle <- air.mle$load #loadings
psi.mle <- air.mle$unique # psi’s
resi.mle <- R-lam.mle%*%t(lam.mle)-diag(psi.mle) #residuals resi.mle
resi.mle
psi.mle
```

PCA is only a linear transformation of data. No assumptions are made
about the form of covariance matrix. In contrast, FA considers
transformations from underlying unobservable factors to observed data.

e)  

-   Weighted least squares

```{r}
air.faw <- factanal(air, factors=2, scores="Bartlett")
air.faw$scores
```

-   Regression approach

```{r}
fit <- factanal(air, 2, rotaion="varimax", scores = "regression")
fit$loadings
fit$scores
```

# Question 3

a)  

-   Correlation matrix S and the sample principal components.

```{r}
stock <- read_excel("stock.xls")
S <- cor(stock)
S
s.s <- eigen(S)
s.s$vectors
s.s$values
```

b)  

-   The proportion and the cumulative proportion of the total variance

```{r}
s.s$values/sum(s.s$values)
cumsum(s.s$values)/sum(s.s$values)
```

c)  

```{r}
pl=princomp(~stock$`Allied Chemical`+stock$`Du Pont`+stock$`Union Carbide`
               +stock$Exxon+stock$Texaco, cor=TRUE, data=stock) 
pl$loadings
plot(pl, type="lines")
```

-   Base on the graph, we should reducing dimension from 7 to 2.
